# Outline of Content & Deliverables

## Works in Progress
  
  [WHAT IS THE BEST WAY TO MAKE/ CODE THIS TABLE? ]
  
```{r library, message=FALSE, warning=FALSE, include=FALSE}
library(DT)
library(tidyverse)
```

```{r create_tbl, include=FALSE}
tbl <- 
  tibble::tribble(
    ~Tool,        ~Tutorial,         ~'Demo Site',          ~Repo,
    #-----------|---------------------|--------------------|-------------------------------------------------
    "Package",  "How to make a package", "NA",              '<a href= "https://github.com/rstudio4edu/testpackage">test package</a>',
    "RMD site", "Basic Fundamentals",    "NA",              "TBA", 
    "RMD site",  "Custom Site Tutorial",   '<a href="https://rstudio4edu.github.io/basic-workshop-site/">Custom Workshop site</a>', '<a href="https://github.com/rstudio4edu/basic-workshop=site">basic-workshop-site</a>',
    "RMD site",  "Custom Site Tutorial",     '<a href="https://rstudio4edu.github.io/basic-course-website/"> Custom Course site</a>', '<a href="https://github.com/rstudio4edu/basic-course-site">basic-course-site</a>',
    "RMD site ",  "Fancy R Markdown Sites",   '<a href="https://rstudio4edu.github.io/fancy-workshop-site/">Fancy Workshop site</a>', '<a href="https://github.com/rstudio4edu/fancy-workshop-site" >fancy-workshop-site</a>',
    "RMD site",   "Fancy R Markdown Sites",    '<a href="https://rstudio4edu.github.io/fancy-course-site"> Fancy Course Site</a>',                    '<a href="https://github.com/rstudio4edu/fancy-course-site">fancy-course-site</a>')
```


```{r use_dt, echo=FALSE}
tbl %>% DT::datatable(escape = FALSE, options=list(dom ="t"))
```

<br>

Now, coding the table a second way, so as not to use HTML explicitly. But now everything is a link.

<br>


```{r, echo = FALSE, message = FALSE}
library(dplyr)
library(knitr)
library(kableExtra)
```

```{r, echo = FALSE}
tbl <- tibble(Tool = c("Package", rep("RMD site", 5)),
              `Tutorial` = c("How to Make a Package",
                                 "Basic Fundamentals", 
                                 rep("Custom Site Tutorial", 2), 
                                 rep("Fancy R Markdown Sites", 2)),
              `Demo site` = c(NA,
                              NA,
                              "Custom Workshop Site",
                              "Custom Course Site",
                              "Fancy Workshop Site", 
                              "Fancy Course Site"),
              `GitHub repo` = c("testpackage",
                                NA,
                                "basic-workshop-site",
                                "basic-course-site",
                                "fancy-workshop-site",
                                "fancy-course-site")
              )

tutorial_url <- c("07-package-tutorial.html",
              "Basic Fundamentals", 
              "r-markdown-custom-site-tutorial.html",
              "r-markdown-custom-site-tutorial",
              "fancy-rmarkdown-sites.html",
              "fancy-rmarkdown-sites.html")

demo_url <- c(NA,
              NA, 
              "https://rstudio4edu.github.io/basic-workshop-site/",
              "https://rstudio4edu.github.io/basic-course-website/",
              "https://rstudio4edu.github.io/fancy-workshop-site/",
              "https://rstudio4edu.github.io/fancy-course-site")

repo_url <- c("https://github.com/rstudio4edu/testpackage",
              NA,
              "https://github.com/rstudio4edu/basic-workshop-site",
              "https://github.com/rstudio4edu/basic-course-site",
              "https://github.com/rstudio4edu/fancy-workshop-site",
              "https://github.com/rstudio4edu/fancy-course-site")

 tbl %>% 
  mutate(
    Tutorial = cell_spec(`Tutorial`, "html", link = tutorial_url),
    `Demo site` = cell_spec(`Demo site`, "html", link = demo_url),
    `GitHub repo` = cell_spec(`GitHub repo`, "html", link = repo_url)) %>%
  kable("html", escape = FALSE) %>%
  kable_styling(bootstrap_options = c("hover")) 


```

## Tutorial Descriptions

### Package

* Most basic package will start right off the bat with data. 

   
### R Markdown Basic Fundamentals

* Alison
* Single page tutorial
* Review syntax
      + See: [items to include](https://github.com/rstudio4edu/rmd-tutorial-workshop/issues/5)    
      
### Custom R Markdown Site
* Applies to both workshops and courses
* Customization within YAML including [these items](https://github.com/rstudio4edu/rmd-tutorial-workshop/issues/5)
* In two parts (chapters?): 
  + How to Build it 
  + What to include in it (could be its own chapter-- or two?): content to be tailored for Workshop use and Courses use
      
### Add-on: How to Build a Fancy R Markdown Site
  + Applies to both workshops and courses
  + Covers anything that would require CSS
     
    
### 