# Outline of Content & Deliverables

## Works in Progress
  
```{r library, message=FALSE, warning=FALSE, include=FALSE}
library(DT)
library(tidyverse)
```

```{r create_tbl, include=FALSE}
tbl <- 
  tibble::tribble(
    ~Tool,        ~Tutorial,         ~'Demo Site',          ~Repo,
    #-----------|---------------------|--------------------|-------------------------------------------------
    "Package",  "How to make a package", "NA",              '<a href= "https://github.com/rstudio4edu/testpackage">test package</a>',
    "RMD site", "Basic Fundamentals",    "NA",              "TBA", 
    "RMD site",  "Custom: Workshop",   '<a href="https://rstudio4edu.github.io/rmd-bare-demo-workshop/">Custom Workshop site</a>', '<a href="https://github.com/rstudio4edu/rmd-bare-demo-workshop">rmd-bare-demo-workshop</a>',
    "RMD site",  "Custom: Course",     '<a href="https://rstudio4edu.github.io/basic-course-website/"> Custom Course site</a>', '<a href="https://github.com/rstudio4edu/basic-course-site">basic-course-site</a>',
    "RMD site ",  "Fancy: Workshop",   '<a href="https://rstudio4edu.github.io/fancy-workshop-site/">Fancy Workshop site</a>', '<a href="https://github.com/rstudio4edu/fancy-workshop-site" >fancy-workshop-site</a>',
    "RMD site",   "Fancy: Course",    "TBA",                    "TBA")
```


```{r use_dt, echo=FALSE}
tbl %>% DT::datatable(escape = FALSE, options=list(dom ="t"))
```

## Tutorial Descriptions

### Package

* Most basic package will start right off the bat with data. 

   
### R Markdown Basic Fundamentals

* Alison
* Single page tutorial
* Review syntax
      + See: [items to include](https://github.com/rstudio4edu/rmd-tutorial-workshop/issues/5)    
      
### Custom R Markdown Site
* Applies to both workshops and courses
* Customization within YAML including [these items](https://github.com/rstudio4edu/rmd-tutorial-workshop/issues/5)
* In two parts (chapters?): 
  + How to Build it 
  + What to include in it (could be its own chapter-- or two?): content to be tailored for Workshop use and Courses use
      
### Add-on: How to Build a Fancy R Markdown Site
  + Applies to both workshops and courses
  + Covers anything that would require CSS
     
    
### 