# (PART) Cookbook: Bookdown book {-}

# Introduction {-#intro-bookdown}

Overview- this is a section filled with recipes for making common types of educational content...


# Make a book

This cookbook is intended for educators who would like to make a "book" to organize their educational content for their audience. Just because we'll be using bookdown doesn't mean that what you intend to make *has* to be an actual book--but because bookdown has a built-in and scrolling table of contents, it's well-suited when you have a lot of content (and many R Markdown files) that you'll need to organize. Consider using bookdown for any long-form tutorial--you can even place an entire course-worth's of materials in a bookdown book. 

Plus, you can also easily:

* Cross-reference
* Number figures, equations, and tables
* Distribute your book as a PDF, an HTML website, etc.

Here, we'll guide you through creating an out-of-the-box bookdown book with basic customization. Then we'll provide suggestions on how to use a book for a workshop or for a course and provide a demo book for each of these use-cases. You can follow along with our demo to make a book of your own from scratch, or you can download the repository containing our demo workshop or course book and modify it as much as you'd like. 


## Is this tutorial for you?

```{r child= "modules/is-this-tutorial-for-you.Rmd"}
```

* How to compile multiple R Markdown files into a Bookdown book
* How to add basic customization through themes using built in options
* How to publish your Bookdown for free via GitHub Pages

## Mental Framework: what are we going to make here? {-}

If you have experience making basic R Markdown sites, then you can think of a Bookdown as an R Markdown site with some additional YAML files (there are THREE total) and some more rigid rules about what can start each document. 


## Part 1: Getting set up {-}

We'll use the following packages:

```{r, eval=FALSE}
#install.packages("bookdown")

#install.packages("usethis")
library(usethis)
```

### Two getting-started workflows:

```{r child = "modules/github-first-project.Rmd"}
```

```{r, eval=FALSE}
bookdown:::bookdown_skeleton(getwd())
```

(note that we use ::: here, with three colons!) to generate important infrasctructure files we need.

Though we recommend a “GitHub first” workflow, it is possible to start with RStudio project wizard to create a new R book project and then move to GitHub:

1. Click *File* > *New Project* > *New Directory*
2. Scroll down and select: *Book Project using bookdown*
3. Then, optionally, to connect to GitHub:

```{r, eval=FALSE}
usethis::use_git()
usethis::use_github() # you have to have a PAT setup
```

### Rendering a skeletal book

Regardless of which workflow you used, you should now have a directory structure with contents like these:
![](INSERT SCREEN SHOT)

These files here are enough to create a skeleton of a book. Let's render it!

:::rstudio-tip
Do not click on *Knit* to render your book! **[WHY?]** If you've already knit, then just clear out the `_book` folder. 
:::

* In the RStudio toolbar menu, click on *Addins* > *Preview Book*

![INSERT SCREEN SHOT of addins]()

But you can also render it these other ways instead:

* 1)  The line below does the exact same thing as *Preview Book*:
```{r, eval=FALSE}
bookdown::render_book('index.Rmd')
```

* 2) Or you can use:
```{r, eval=FALSE}
bookdown::serve_book()
```



### Three kinds of YAMLs

In RStudio navigate to... 

`index.Rmd` is going to be Chapter 1.

## Part 2: Create your content {-}


### Structure

#### Files to change
#### What not to touch

#### Figures and Tables
#### Cross References
#### Citations
#### References

## Part 3: Dress it up {-}

### Highlight options
Possible styles are `default`, `tango`, `pygments`, `kate`, `monochrome`, `espresso`, `zenburn`, and `haddock`.
### Chapters or...? {-}
### Numbered or Not {-}



## Part 4: Put your book out there! {-}

### Online
### As a PDF
### As....


## Workflow for maintenance {-}

Once you have the website in place, it’s likely that you will want to update it periodically or add new pages. Here’s what that workflow looks like for us:  

1. Open my existing site RStudio project
2. Go to *Addins* > *Preview Book* to render the book
3. 
4. 
5. 

## Gotchas to watch out for

(Or call this section, Notes for using your book once it's made)

- We've already covered not knitting when rendering your book.
- All chapter `.Rmd` files MUST begin with # Level 1 Header no matter what. Any content before that will not appear.
- Making sure your book contents are put together in the right order: For this to happen automatically, you would need to create .Rmd file names that start with 01, 02, etc. This seems fine and dandy, but think of how annoying it would be if you decided to switch around the order of all your chapters--you would have to rename multiple files (ugh!). So, it is recommended to instead go to the `_bookdown.yml` and set the chapter order there. This is kind of like the bookdown builder's instruction manual for chapter order. Swap around your chapters as much as you'd like in this single file and be done.
- In `_bookdown.yml`, when you fill out `chapter_name`: What follows should include a space after it, within the quotes!
- Explain the never-ending loop of book-rendering [Would be good to get some more context into what sorts of things trigger this behavior.] Common problem! Exit out of this by going to *Session* > *Terminate R...*.
- re-explain search here https://github.com/rstudio/bookdown/issues/216

## Bookdown resources
[https://bookdown.org/yihui/bookdown/](https://bookdown.org/yihui/bookdown/)


# Make a book for a workshop

# Make a book for a course

# Make it fancier

Once you’ve built the basic structure for your book, if you’d like to customize it even further, we can do a little bit of HTML and CSS work to add-on some fancy features.  

## Anatomy of a fancy book 
Here's where we're starting: 

[SCREENSHOT OF PLAIN BOOK]

And here's where we're going:

[SCREENSHOT OF FANCY BOOK, ANNOTATED]

**Link back to CSS Crash Course** 

## Add Google Fonts {-}
## Add hero image {-}
## Table of contents
### Hierarchy styling
### Background
## Add a logo
## Sidebar Divs
## Drop Caps
## Copy to Clipboard




