# (PART) Cookbook: Bookdown book {-}

# Introduction {-#intro-bookdown}

Overview- this is a section filled with recipes for making common types of educational content...


# Make a book

This cookbook is intended for educators who would like to make a "book" to organize their educational content for their learners. Just because we'll be using bookdown doesn't mean that what you intend to make *has* to be an actual book--but because bookdown has a built-in and scrolling table of contents, it's well-suited when you have a lot of content (and many R Markdown files) that you'll need to organize. Consider using bookdown for any long-form tutorial--you can even place an entire course-worth's of materials in a bookdown book. 

Plus, you can also easily:

* Cross-reference
* Number figures, equations, and tables
* Distribute your book as a PDF, an HTML website, etc.

Here, we'll guide you through creating an out-of-the-box bookdown book with basic customization. Then we'll provide suggestions on how to use a book for a workshop or for a course and provide a demo book for each of these use-cases. You can follow along with our demo to make a book of your own from scratch, or you can download the repository containing our demo workshop or course book and modify it as much as you'd like. 

## Mental Framework: what are we going to make here? {-}

If you have experience making basic R Markdown sites, then you can think of a Bookdown as an R Markdown site with some additional YAML files and some more rigid rules about what can start each document. 



## Part 1: Getting set up {-}

We'll use the following packages:

```{r, eval=FALSE}
#install.packages("bookdown")

#install.packages("usethis")
library(usethis)
```

### Two getting-started workflows:

```{r child = modules/github-first-project}
```

```
bookdown:::bookdown_skeleton(getwd())
```

(note that we use ::: here, with three colons!) to generate important infrasctructure files we need.

Though we recommend a “GitHub first” workflow, it is possible to start with RStudio project wizard to create a new R book project and then move to GitHub:

1. Click *File* > *New Project* > *New Directory*
2. Scroll down and select: *Book Project using bookdown*

## Part 2: Create your content {-}

## Part 3: Dress it up {-}

### Chapters or...? {-}
### Numbered or Not {-}

## Part 4: Put your book online! {-}

## Workflow for maintenance {-}

Once you have the website in place, it’s likely that you will want to update it periodically or add new pages. Here’s what that workflow looks like for me:  

1. Open my existing site RStudio project
2. 
3. 
4. 
5. 

## Gotchas to watch out for

(Or call this section, Notes for using your book once it's made)

- re-explain search here https://github.com/rstudio/bookdown/issues/216

## Bookdown resources
[https://bookdown.org/yihui/bookdown/](https://bookdown.org/yihui/bookdown/)


# Make a book for a workshop

# Make a book for a course

# Make it fancier

Once you’ve built the basic structure for your book, if you’d like to customize it even further, we can do a little bit of HTML and CSS work to add-on some fancy features.  

## Anatomy of a fancy book 
Here's where we're starting: 

[SCREENSHOT OF PLAIN BOOK]

And here's where we're going:

[SCREENSHOT OF FANCY BOOK, ANNOTATED]

**Link back to CSS Crash Course** 

## Add Google Fonts {-}
## Add hero image {-}
## Table of contents
### Hierarchy styling
### Background
## Add a logo
## Sidebar Divs
## Drop Caps
## Copy to Clipboard




