# (PART\*) Teaching with lessons {-}

# Introduction {-#intro-lessons}

Note: not teaching others *how to use* R Markdown, but teaching yourself tips/tricks for how to teach others *with* R Markdown.

This will be for static edu materials (as opposed to interactive- i.e., learnR)

# Do you need R Markdown?

Maybe not, can do .R scripts, knitr::spin to GitHub

# R Markdown in practice

examples:

+ https://bbc.github.io/rcookbook/

why, how

also include how to leave invisible notes to yourself `<!--- your comment --->`

## R Markdown Workflows

+ Using RStudio projects (why)
    + https://www.tidyverse.org/articles/2017/12/workflow-vs-script/
    + in RStudio, from git adds .gitignore for you
+ file paths using `here::here()`
    + https://malco.io/2018/11/05/why-should-i-use-the-here-package/

# Design for teaching

## Formatting code

Useful patterns:

+ Just show me the code

+ Just show me the output

+ Just show me the plots

+ [verbatim code chunks](https://yihui.name/en/2017/11/knitr-verbatim-code-chunk/)
+ `knitr` code chunk options (like eval vs echo vs [include](https://yihui.name/en/2017/11/knitr-include-false/))
+ `knitr` [ref.label](https://yihui.name/knitr/demo/reference/) to reuse code chunks
+ the `fig.path` thing: https://yihui.name/knitr/options/#plots (see also: name your code chunks!)
+ making sure that your package warnings/messages don't print!
+ how to control numbers of digits that print 
+ control significance stars

```{r}
options(
  show.signif.stars = FALSE,     
  digits = 2
  )
```


## Styling output

HTML output options like floating toc, numbering, code folding, code download (requires theme)

- [ ] [tabsets](https://bookdown.org/yihui/rmarkdown/html-document.html#tabbed-sections)
         - think about good use cases
- [ ] [dropdown in tabsets](https://github.com/rstudio/rmarkdown/pull/1405)
- [ ] [code folding](https://rmarkdown-docs.netlify.com/reference/html_document.html)
- [ ] spacing with `<br>`
- [ ] highlight
- [ ] themes (plus `type: inverse`) 
- [ ] [toc / toc floating](https://rmarkdown-docs.netlify.com/reference/html_document.html) 

- [breadcrumbs](https://bootswatch.com/flatly/)

# Sharing 

## Share data

For html output formats only:

1. `xfun::embed_file()` [link](https://yihui.name/en/2018/07/embed-file/)

1. DT buttons [link](https://rstudio.github.io/DT/003-tabletools-buttons.html)

```{r eval = FALSE}
library(DT)
datatable(payroll, extensions = 'Buttons', options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
  )
)
```


## Share source code

+ first, know the knitr code chunk options, and remember about code folding.

+ GitHub to view source (but not very readable, see next section)

- [ ] [code download](https://rmarkdown-docs.netlify.com/reference/html_document.html)

## Share readable output

+ .R -> GitHub

  ```yaml
  #' ---
  #' output: github_document
  #' ---
  ```

  + https://github.com/apreshill/teachthat/blob/master/gather/make-data-to-gather.R

+ Rmd -> GitHub

   + https://rmarkdown.rstudio.com/docs/reference/github_document.html
   + https://github.com/jennybc/lotr-tidy

+ GitHub + ghpages

Can also export to pdf as attachments
      
   
# Troubleshooting

[Mine's troubleshooting tips:](https://mine-cetinkaya-rundel.github.io/teach-data-sci-icots2018/01-03-rmarkdown/01-03-rmarkdown.html#30)



+ Make sure RStudio and the rmarkdown package (and its dependencies) are up-to-date.

+ Set a global option for error = TRUE (or for a given chunk) so that your document renders even when there are errors.

+ Don’t try to change working directory within an R Markdown document. (If you do still decide to use setwd in a code chunk, beware that the new working directory will only apply to that specific code chunk, and any following code chunks will revert back to use the original working directory.)

Also see: 

+ https://rmd4sci.njtierney.com/common-problems-with-rmarkdown-and-some-solutions.html
+ https://happygitwithr.com/rmd-test-drive.html#rmd-troubleshooting
+ switch latex engines (do a screenshot here)
+ consider printing out your `sessionInfo()` [for pro-level, use DT]

```{r}
DT::datatable(devtools::session_info()$packages)
```

